---
title: "Drake Lyrics Analysis"
author: "avril"
date: "2025-10-29"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(tidyverse)
library(stringr)
library(tidytext)
library(kableExtra)
library(tm)
library(wordcloud)
library(SnowballC)
```

# Intro

I started with a dataset from [[**Kaggle**]{.underline}](https://www.kaggle.com/datasets/juicobowley/drake-lyrics?resource=download) containing all of Drake's lyrics at the time of its creation. In this project, I perform textual normalization on song lyrics, evaluate word frequency distributions, analyze key terms across albums, and investigate the relationship between lyric themes (specifically the word "love") and overall track popularity.

```{r}
#cleaning up the data
drake_combined=c()
for(i in 1:11){
  d=read.csv(paste0("drake_data_",i,".csv"))
  drake_combined=rbind(drake_combined,d)
}
drake_combined$lyrics=str_to_lower(drake_combined$lyrics)
drake_combined=mutate(drake_combined,
                      lyrics = gsub("[&\\:\\,\\.\\?\\![:digit:]\\\n]", "",lyrics))
```

# Top 10 words

This table displays the top 10 most frequently occuring words.

```{r}
lyrics_only=drake_combined$lyrics
lyrics_words=str_split(lyrics_only," ", simplify=TRUE)
lyrics_words_vec=as.vector(lyrics_words)
lyrics_words_vec=lyrics_words_vec[lyrics_words_vec!=""]
my.table=as.data.frame(table(lyrics_words_vec))
my.sort=arrange(my.table,desc(Freq))
colnames(my.sort)=c("Word","Frequency")
kbl(my.sort[1:15,])
```

However, the words are mostly filler words, so here I cleaned up the data a bit to make a more informative table.

```{r}
freq_lyrics <- lyrics_words_vec[!lyrics_words_vec %in% stop_words$word]
freq_table <- as.data.frame(table(freq_lyrics))
freq_sorted <- arrange(freq_table,desc(Freq))
colnames(freq_sorted)=c("Word","Frequency")
kbl(freq_sorted[1:15,])
```

Now, a word cloud. I used [[***this***]{.underline}](https://lukesingham.com/how-to-make-a-word-cloud-using-r/) very helpful tutorial!

```{r}
wordcloud(freq_sorted$Word, freq_sorted$Frequency, scale=c(5,0.5) 
        , 100   
        , random.order=FALSE 
        , rot.per=0.35       
        , use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))
```

Definitely some interesting results!

# Occurence of words across albums

Here I choose five words and count their occurrences across albums

```{r}
my.albums=unique(c(drake_combined$album))
table=c()
my.words=c("love","drake","respect","money","trust")
for(j in 1:length(my.albums)){
  my.sub=drake_combined[drake_combined$album==my.albums[j],]
  counts=vector(length=5)
  for(i in 1:5){
    counts[i]=sum(str_count(my.sub$lyrics,my.words[i]))
  }
  table=rbind(table,counts)
}
table=as.data.frame(table)
colnames(table)=my.words
table=cbind(my.albums,table)
kable_classic(kbl(table,row.names=FALSE))
```

Now in barplot form

```{r}
table_long=pivot_longer(table,cols=my.words,names_to="word",values_to="count")
for(i in 1:5){
  sub=table_long[table_long$word==my.words[i],]
  myplot=ggplot(sub)+geom_bar(aes(x=my.albums,
                                  y=count,fill=my.albums),stat="identity") + labs(title=paste("Counts of the word:",
                                                                                              my.words[i]),x="Album name",y="Count") + theme(legend.position = "none")+ coord_flip() 
  print(myplot)
}
```

Here I track the success of tracks by use of the word "love"

```{r}
drake_combined$love=str_count(drake_combined$lyrics,"love")
view_number=c()
for(i in 1:length(drake_combined$track_views)){
  if(is.na(drake_combined$track_views[i])){
    view_number=c(view_number,NA)
  }
  else if(str_detect(drake_combined$track_views[i],"K")==TRUE)
  {
    view_number[i]=parse_number(drake_combined$track_views[i])*1000
  }
  else if(str_detect(drake_combined$track_view[i],"M")==TRUE)
  {
    view_number[i]=parse_number(drake_combined$track_views[i])*1000000
  }
  else
  {
    view_number[i]=parse_number(drake_combined$track_views)
  }
}
drake_combined$track_views=view_number
ggplot(drake_combined) + geom_point(aes(x=love,
                                        y=track_views,color=album)) + labs(title="Uses of the word 'Love' by track views", x="Frequency",y="Track views")
```

We can see that, in general, his tracks that did the best didn't use the word "love" too much.

# Conclusion/Reflections

We definitely saw some pretty interesting results overall. We saw that tracks with high view counts (\> 50M views) usually feature lower frequencies of the word "love". Also, the most frequently used words were a bit surprising to me, I was definitely not expecting "love" to be the top word!

